---
title: "Severe weather events"
author: "Koji"
date: "2018/7/9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

> Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences.

## Data

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

[Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) [Mirror Site]

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.

## Data Processing

> There should be a section titled Data Processing which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the cache=TRUE option for certain code chunks.

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
#download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", destfile = "repdata-data-StormData.csv.bz2")
```

```{r cache=TRUE}
stormData <- read.csv("repdata-data-StormData.csv.bz2", header = TRUE, stringsAsFactors = FALSE)
```

```{r}
str(stormData)
```

```{r}
alphabeticalExp <- function(vec) {
        mapply(function(alphabet) {
                ifelse(alphabet == "",
                       10^0, # 1
                       switch(tolower(alphabet),
                              "h" = 10^2, # hundred
                              "k" = 10^3, # thouzand
                              "m" = 10^6, # million
                              "b" = 10^9, # billion
                              "+" = 10^0, # 1
                              "-" = 10^0, # 1
                              "?" = 10^0, # 1
                              " " = 10^0, # 1
                              10^as.integer(alphabet)))
        }, vec)
}
```

```{r}
humanImpactTop10 <- stormData %>%
        select(EVTYPE, FATALITIES, INJURIES) %>%
        mutate(PERSONS = FATALITIES + INJURIES) %>%
        select(EVTYPE, PERSONS) %>%
        group_by(EVTYPE) %>%
        summarize_all(sum) %>%
        arrange(desc(PERSONS)) %>%
        top_n(10, PERSONS)

humanImpactTop10
```

```{r}
economicImpactTop10 <- stormData %>%
        select(EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP) %>%
        mutate(DAMAGE = PROPDMG * alphabeticalExp(PROPDMGEXP) +
                        CROPDMG * alphabeticalExp(CROPDMGEXP)) %>%
        select(EVTYPE, DAMAGE) %>%
        filter(!is.na(DAMAGE)) %>%
        group_by(EVTYPE) %>%
        summarize_all(sum) %>%
        arrange(desc(DAMAGE)) %>%
        top_n(10, DAMAGE)

economicImpactTop10
```


## Results

### Across the United States, which types of events are most harmful with respect to population health?

```{r}
ggplot(data = humanImpactTop10, aes(x = reorder(EVTYPE, -PERSONS), y = PERSONS)) +
        geom_bar(stat = "identity", position = position_dodge(), fill = "orangered") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        ggtitle("Human Impact")
```

### Across the United States, which types of events have the greatest economic consequences?
```{r}
ggplot(data = economicImpactTop10, aes(x = reorder(EVTYPE, -DAMAGE), y = DAMAGE)) +
        geom_bar(stat = "identity", position = position_dodge(), fill = "orangered") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        ggtitle("Economic Damage")
```


> There should be a section titled Results in which your results are presented.

> There should be a section titled Results in which your results are presented.
> You may have other sections in your analysis, but Data Processing and Results are required.
> The analysis document must have at least one figure containing a plot.
> Your analysis must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total.
> You must show all your code for the work in your analysis document. This may make the document a bit verbose, but that is okay. In general, you should ensure that echo=TRUE for every code chunk (this is the default setting in knitr).
